Association between Breast Cancer and Hepatitis C Virus with Age Controlling

### 2022-11-14 manuscript/manuscript.tex
equations

\begin{abstract}
The Kolmogorov--Smirnov (KS) test is one of the most popular goodness-of-fit
tests for comparing a sample with a hypothesized continuous distribution.
Nevertheless, it has often been misused.

@@

Specifically, this paper will determine if there is an association between testing positive for hepatitis C virus
and breast cancer while controlling for age. Both Hepatitis C and breast cancer are extremely prevalent in 
Egypt \citep{Hussein2021high}. Very recently on October 6th, Egypt and Qatar agreed to collaborate in the 
health sector and use each 
other's expertise. The health ministers specifically mentioned research in Hepatitis C and breast cancer
as a area of interest \citep{arham2022egypt}. Thus, research surrounding these diseases is becoming more
important in the field.

@@

The data used in the paper is from a retrospective case control study 
conducted in 2020 \citep{2020association}. The data includes 405 subjects 
as part of the study group, all having been treated for breast cancer in the 
past 10 years. The second group consists of 145 adult females from a 
governorate in Egypt, who all participated in a cross-sectional study 
from 2015-2017. This data can be put into a 2x2 contingency table
for statistical analysis.

To test the association of breast cancer and Hepatitis C, one can use
the chi-square or Fisher's Exact test, depending on the
conditions from the sample size \citep{warner2013testing}. Fisher's
exact test is useful for when the normality assumption is violated 
and the expected values of the 2x2 table are too small. The test uses 
the hypergeometric distribution to test if the probabilities are
the same between the two groups. Thus, we can determine if there is 
more of a risk of breast cancer for those with seropositive Hep C. As
 a result, we can determine if women who chronically test positive should 
 be tested more frequently because they are at a higher risk.

Figure~\ref{fig:formula} shows the probability of the original table occuring.

\begin{figure}[tbp]
  \centering
  \includegraphics[width=8cm]{formula.png}
  \caption{Hypergeometric pmf.}
  \label{fig:formula}
\end{figure}

\section{Application}
\label{sec:app}

@@

\begin{tabular}{ | m{4cm} | m{4cm}| m{4cm} | m{2cm} | }
  \hline
    & HCV Seropositive & HCV Seronegative & Total\\ 
  \hline
  Breast Cancer & 88 & 317 & 405 \\ 
  \hline
  No Cancer Diagnosis & 15 & 130 & 145 \\ 
  \hline
  Total & 103 & 447 & 550 \\ 
  \hline
\end{tabular}

@@

\begin{tabular}{ | m{4cm} | m{4cm}| m{4cm} | m{2cm} | }
  \hline
    & HCV Seropositive & HCV Seronegative & Total\\ 
  \hline
  Breast Cancer & 17 & 110 & 127 \\ 
  \hline
  No Cancer Diagnosis & 2 & 91 & 93 \\ 
  \hline
  Total & 19 & 201 & 220 \\ 
  \hline
\end{tabular}

@@

\begin{tabular}{ | m{4cm} | m{4cm}| m{4cm} | m{2cm} | }
  \hline
    & HCV Seropositive & HCV Seronegative & Total\\ 
  \hline
  Breast Cancer & 71 & 207 & 278 \\ 
  \hline
  No Cancer Diagnosis & 13 & 39 & 52 \\ 
  \hline
  Total & 84 & 246 & 330 \\ 
  \hline
\end{tabular}

>>>>

\begin{abstract}
In Egypt Hepatitus C virus infection is a common problem that often impedes treatment
of Breast Cancer in female patients. While previous work has analyze an associaiton 
between HCV seropositivity and breast cancer diagnosis, this paper will focus on the 
potential confounding of age on the association. We will use Fisher's Exact Test to
estimate the crude estimate of the association and the Mantel-Haenszel method to 
stratify the data between ages. The possible effect of age in the association can
lead to improved care for patients.

@@

Specifically, this paper will determine if there is an association between testing positive for hepatitis C virus
and breast cancer while controlling for age in Egyptian populations. Both Hepatitis C and breast 
cancer are extremely prevalent in Egypt \citep{Hussein2021high}. Very recently on October 6th, 
Egypt and Qatar agreed to collaborate in the health sector and use each 
other's expertise. The health ministers specifically mentioned research in Hepatitis C and breast cancer
as a area of interest \citep{arham2022egypt}. Thus, research surrounding these diseases is becoming more
important in the field.

@@

The data used in this paper is from a 2020 retrospective case-control 
study based on Egyptian female populations \citep{2020association},
sponsered by Mansoura University in Egypt.

The study group consists of 405 patients treated at the Oncology Center 
- Mansoura University in the past 10 years. In order to be included, 
invasive breast cancer must be biopsy-proven. Patients with unknown 
viral marker status, multiple cancer diagnoses and virus-unrelated 
hepatic pathology were all excluded.

The control group consists of 
data from a previous study, conducted from 2015-2017. There are 145 
females from the same geographic distribution, all with no previous 
cancer diagnosis. In both groups, all patients are above the age of 
18. For this paper, patients are stratified based on age- older patients 
are considered 45 years or above and younger patients are below 45 years. 
The data is summarized in the following 2x2 contingency tables.

@@

\begin{tabular}{ | m{4cm} | m{4cm}| m{4cm} | m{2cm} | }
  \hline
    & Breast Cancer & No Cancer Diagnosis & Total\\ 
  \hline
  HCV Seropositive & 88 & 15 & 103 \\ 
  \hline
  HCV Seronegative & 317 & 130 & 447 \\ 
  \hline
  Total & 405 & 145 & 550 \\ 
  \hline
\end{tabular}

@@

\begin{tabular}{ | m{4cm} | m{4cm}| m{4cm} | m{2cm} | }
  \hline
    & Breast Cancer & No Cancer Diagnosis & Total\\ 
  \hline
  HCV Seropositive & 17 & 2 & 19 \\ 
  \hline
  HCV Seronegative & 110 & 91 & 201 \\ 
  \hline
  Total & 127 & 93 & 220 \\ 
  \hline
\end{tabular}

@@

\begin{tabular}{ | m{4cm} | m{4cm}| m{4cm} | m{2cm} | }
  \hline
    & Breast Cancer & No Cancer Diagnosos & Total\\ 
  \hline
  HCV Seropositive & 71 & 13 & 84 \\ 
  \hline
  HCV Seronegative & 207 & 39 & 246 \\ 
  \hline
  Total & 278 & 52 & 330 \\ 
  \hline
\end{tabular}

\begin{figure}[tbp]
  \centering
  \includegraphics[width=8cm]{table.png}
  \caption{2x2 Contingency Table}
  \label{fig:table}
\end{figure}

\begin{figure}[tbp]
  \centering
  \includegraphics[width=8cm]{formula.png}
  \caption{Hypergeometric pmf}
  \label{fig:formula}
\end{figure}

\begin{figure}[tbp]
  \centering
  \includegraphics[width=12cm]{2x2 stratified table.png}
  \caption{2x2 Contingency Table with Stratified Data}
  \label{fig:strat}
\end{figure}

\begin{figure}[tbp]
  \centering
  \includegraphics[width=8cm]{breslow day test stat.png}
  \caption{Breslow Day Test Statistic}
  \label{fig:BD}
\end{figure}

\begin{figure}[tbp]
  \centering
  \includegraphics[width=8cm]{MH Test Stat.png}
  \caption{Mantel-Haenszel Test Statistic}
  \label{fig:MH}
\end{figure}

To test the association of breast cancer and Hepatitis C seropositivity, one can use
the chi-square or Fisher's Exact test, depending on the
conditions from the sample size \citep{warner2013testing}. Fisher's
exact test is useful for when the normality assumption is violated 
and the expected values of the 2x2 table are too small. The test uses 
the hypergeometric distribution to test if the probabilities are
the same between the two groups. Thus, we can determine if there is 
more of a risk of breast cancer for those with seropositive Hep C.
In the data section, the data was summarized into 2x2 contingency tables.
The generalzied table is shown in Figure~\ref{fig:table}.
The first step is calculate the probability of the original table occuring,
using the hypergeometric pdf. Figure~\ref{fig:formula} shows the formula used.

Based on the alternative hypothesis, you then calculate the probability of 
all more extreme tables based on the a value. In this paper, p1 is the
probability of having breast cancer given that the patient is seropositive
and p2 is the probability of having breat cancer given that the patient
is seronegative. Then, we want to determine if there is an association,
specifically if the risk of having breast cancer is higher for those who
are seropositive. The alternative hypothesis is that p1 is greater than p2-
which will be tested in the application section.

After calculating the crude estimate, a Breslow-Day test will test for 
homogeneity of odds ratios between the age stratified data. The generalized
form of stratified 2x2 contingency tables is shown in Figure~\ref{fig:strat}.
The Breslow-Day test statistic is shown in Figure~\ref{fig:BD}, which will
be computed using SAS.

Based on the results of the Breslow Day test for homogeneity of odds ratios,
the Mantel-Haenszel test will compute a common odds ratio and test if there 
is an association between seropositivity and breast cancer while controlling 
for age. The Mantel-Haenszel test statistic is shown in Figure~\ref{fig:MH}
where the expected value of A is 
\begin{equation}
  E(A_{i})=\frac{M_{i1}R_{1i}}{N_{i}} 
\end{equation}
and the variance of A is 
\begin{equation}
  V(A_{i})=\frac{R_{1i}R_{0i}M_{1i}M_{0i}}{N_i^2(N_{i}-1)}. 
\end{equation}

\section{Application}
\label{sec:app}
-
